Compositional Certified Resource Bounds\ifshort{\vspace{-.15cm}}{}

### 2014-11-08 paper/main.tex
add logical state examples

\caption{Example derivation using amortized reasoning.  We write
  $\#_1(a)$ for $\# \{ i \mid 0 \le i < N \land a[i] = 1 \}$ and use the
  tick metric that assigns cost $n$ to the statement \code{tick(n)} and
  0 to all other operations.
  }
\label{fig:xmplinc}
\end{figure}
%
\iffull{
}{}
We can then use the logic to prove a concrete bound for a given resource
metric, or a bound that is parametric in a set of resource metrics.
For example, we can derive the following quantitative
triple for a binary-search function \code{bsearch} that holds for all
stack metrics $M$.
%
\begin{center}
${ \{ (Z = \log_2(h-\ell)) + Z{\cdot}M_\code{bsearch} \} }$
\code{bsearch(x,l,h)}
${ \{ Z{\cdot}M_\code{bsearch} \} }$
\end{center}
%
A stack metric assigns cost $0$ to all evaluation steps except
function calls.  Before a call $f(x)$, we consume $M_f > 0$ resources
and after the call we return $M_f$ resources that can be used in
subsequent function calls.  So $M_f$ corresponds to the stack-frame
size of the function $f$.  In the pre- and postcondition, the logical
variable $Z$ is used to relate the size of the input with the
potential that is left after the function call.  The logical part $Z =
\log_2(h-\ell)$ of the precondition is $0$ if $Z =
\log_2(\state_0(h)-\state_0(\ell))$ in the initial state $\state_0$ and
$\infty$ otherwise.

>>>>

\caption{Example derivation using amortized reasoning.  We write
  $\#_1(a)$ for $\# \{ i \mid 0 \le i < N \land a[i] = 1 \}$ and use the
  tick metric that assigns cost $n$ to the statement \code{tick(n)} and
  0 to all other operations.
  }
\label{fig:xmplinc}
\end{figure}
%
\iffull{
}{}
\begin{figure}
\center
\begin{minipage}[b]{0.8\linewidth}
\begin{lstlisting}
${\color{blue}%
  \mbox{logical state invariant } \{ \code{na} = \#_1(a)} \}$
while (k > 0) {
  x=0;
  while (x < N && a[x] == 1 && na > 0) {
    a[x]=0;
    na--; tick(1);
    x++; }
  if (x < N) {
    a[x]=1;
    na++; tick(1); }
  k--;
}

\caption{Example of assisted bound derivation using logical state.  We write
  $\#_1(a)$ for $\# \{ i \mid 0 \le i < N \land a[i] = 1 \}$ and use the
  tick metric that assigns cost $n$ to the statement \code{tick(n)} and
  0 to all other operations.  The bound derived automatically is
  $2|[0,\code{k}]| + |[0,\code{na}]|$.
  }
\label{fig:xmplincaux}
\end{figure}
%
\begin{figure}
\begin{lstlisting}
${\color{blue}%
  \mbox{logical state invariant }
  \{ \code{lg} \ge \log_2(h-l) \cdot M_\code{bsearch} \} }$
bsearch(x,l,h,lg) {
  if (h-l > 1 && lg > 0) {
    m = h + (h-l)/2;
    lg--; if (a[m]>x) h=m; else l=m;
    tick($M_\code{bsearch}$);
    l = bsearch(x,l,h,lg);
    tick($-M_\code{bsearch}$);
  } else return l;
}
\end{lstlisting}
\caption{Example of assisted bound derivation using logical state.
  We use the tick metric to model the stack consumption of the program.
  Before a function call to \code{bsearch}, a tick statement pays for
  the space used by stack frame of the function.  The space is recoverd
  after the call.
  The bound derived automatically is $|[0,\code{lg}]|$.
  }
\label{fig:xmplbsaux}
\end{figure}
%
\iffull{
}{}
We can then use the logic to prove a concrete bound for a given resource
metric, or a bound that is parametric in a set of resource metrics.
For example, we can derive the following quantitative
triple for a binary-search function \code{bsearch} that holds for all
stack metrics $M$.
%
\begin{center}
${ \{ (Z = \log_2(h-\ell)) + Z{\cdot}M_\code{bsearch} \} }$
\code{bsearch(x,l,h)}
${ \{ Z{\cdot}M_\code{bsearch} \} }$
\end{center}
%
A stack metric assigns cost $0$ to all evaluation steps except
function calls.  Before a call $f(x)$, we consume $M_f > 0$ resources
and after the call we return $M_f$ resources that can be used in
subsequent function calls.  So $M_f$ corresponds to the stack-frame
size of the function $f$.  In the pre- and postcondition, the logical
variable $Z$ is used to relate the size of the input with the
potential that is left after the function call.  The logical part $Z =
\log_2(h-\ell)$ of the precondition is $0$ if $Z =
\log_2(\state_0(h)-\state_0(\ell))$ in the initial state $\state_0$ and
$\infty$ otherwise.
